Issues with Dynamical Measurements Across the Disk

### 2019-02-25 paper/ms.tex
moved table 3 earlier and removed redundant discussion in Section 4

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
used in this work. Cosmological zoom-ins allow one to simulate a selected
region at high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm\ for
shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{available at
\url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi\ contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
and \mm , respectively.

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about 40 pc (3--4 times the
cold gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, and indeed the gas and stellar scale heights of
the simulated galaxies are equal to or larger than the MW's. However, in this
work we are considering azimuthal variations in the midplane \emph{as defined
by the mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

@@

\section{Discussion} \label{sec:discussion}
We have shown that, in simulation, the ``local midplane'' can have strong
azimuthal dependence. The implications of this finding are not clear, but
clearly demonstrate that assuming our local midplane is the same as the global
midplane is insufficient for extracting the most dynamical information from
the {\em Gaia} mission.

@@

\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
simulations of Milky Way-mass galaxies are not perfect representations of the
Milky Way, and do not reproduce every available observation of the Milky Way.
For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
closer to that of M31 \citep{2018arXiv180610564S}.

In order to assess the applicability of our work to the real Milky Way, we
consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
and \mm\ and the current best estimates for the Milky Way. One noticeable
feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
This would seem to suggest that the midplane estimates from
Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.

\begin{deluxetable}{ccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{a} & 300 & 900 \\
\mi\tablenotemark{b} & 480 & 2000 \\
\mf\tablenotemark{b} & 440 & 1280 \\
\mm\tablenotemark{b} & 290 & 1030 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}

\tablenotetext{a}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
\tablenotetext{b}{\citet{2018arXiv180610564S}}
\end{deluxetable}

Another important aspect to consider when comparing the simulated galaxies to
the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
incorporate merger trees?}

>>>>

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
used in this work. Cosmological zoom-ins allow one to simulate a selected
region at high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm\ for
shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{at
\url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi\ contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
and \mm , respectively. These values are summarized in Table \ref{tab:scale_height}.

\begin{deluxetable*}{cccccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{cold\tablenotemark{a} gas disk} & \colhead{stellar thin disk} & \colhead{stellar thick disk} & \colhead{cold gas softening length} & \colhead{stellar softening length} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{b} & 40 & 300 & 900 & \nodata & \nodata \\
\mi\tablenotemark{c} & 800\tablenotemark{d} & 480 & 2000 & 12.4 & 11.2 \\
\mf\tablenotemark{c} & 360 & 440 & 1280 & 12.5 & 11.2 \\
\mm\tablenotemark{c} & 250 & 290 & 1030 &10.3 & 11.2 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}
\tablenotetext{a}{$T<1000$ K}
\tablenotetext{b}{\citet{2008ApJ...673..864J,2016ARAA..54..529B}}
\tablenotetext{c}{\citet{2018arXiv180610564S}}
\tablenotetext{d}{The azimuthally averaged gas vertical density profile in m12i is nearly constant to this height, though individual regions show smaller scale heights and dense clouds.}
\end{deluxetable*}

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about 40 pc (3--4 times the
cold gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, especially the cold gas. Indeed, the stellar scale heights of
the simulated galaxies are equal to or larger than the MW's while the gas scale heights are significantly larger (although the proper basis comparison is less clear in the case of the gas; the quoted value for the Milky Way comes from studies of high-mass star-forming regions). As noted above, the midplanes defined by gas and stars can be tilted with respect to one another as well, precluding extending the precision of the gas midplane definition to the stellar component. In this
work we specifically consider azimuthal variations in the midplane \emph{as defined
by the stellar mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

@@

\section{Discussion} \label{sec:discussion}
We have used high-resolution cosmologial simulations to illustrate that the ``local midplane'' defined by stellar density can have a strong azimuthal dependence, as a natural consequence of the non-axisymmetry of the Galactic disk at small scales. While this is not in itself surprising or new, we also demonstrate that extending the coordinate system established by our local midplane to a global axisymmetric coordinate system spanning the entirety of the Galactic disk introduces systematic error in computations of the actions under this symmetry assumption, when starting from the present-day positions and velocities of stars as measured, for example, by \emph{Gaia}.

@@

%%% this stuff is now in section 3 mostly
%\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
%As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
%simulations of Milky Way-mass galaxies are not perfect representations of the
%Milky Way, and do not reproduce every available observation of the Milky Way.
%For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
%closer to that of M31 \citep{2018arXiv180610564S}.
%
%In order to assess the applicability of our work to the real Milky Way, we
%consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
%Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
%and \mm\ and the current best estimates for the Milky Way. One noticeable
%feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
%\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
%This would seem to suggest that the midplane estimates from
%Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.

%Another important aspect to consider when comparing the simulated galaxies to
%the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
%both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
%had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
%incorporate merger trees?}
